---
title: "Analysis of first visit to ERVITAL"
output: 
        html_document:
          toc: true
author: "Simão Abreu"
date: "2020-12-10"
---

```{r setup,echo=FALSE}
knitr::opts_chunk$set(cache = T,fig.align = "center",echo = F,cache.lazy = FALSE)
```

```{r libraryLoad,include=FALSE,cache = FALSE}
hrm::hrmAttach()
library(patchwork)
library(ggthemes)
library(DT)
```



```{r loadData,echo=FALSE}
workflowData <- read_rds("salvia.rds")
binalysis <- workflowData %>%
                        resultsProcessing()
analysis <- workflowData %>%
                        resultsAnalysis()
assignment <- workflowData %>%
      resultsAnnotation()
```
    
## Overview
      
**Technique:** `metaboWorkflows` v0.8.5 workflow - FIE-HRMS fingerprinting

**Undertaken:** Thu Dec 10 15:46:45 2020

**Flags:** spectralBin, detectBatchDiff, detectMissInjections, preTreat, MFassignment, reduceIsotopes, modelling, correlations

## Spectral Binning
      
### Overview
      
**Technique:** FIE-HRMS Spectral Binning

```{r binalysisOverview,echo = FALSE}
print(binalysis)
```

### Parameters

```{r binParamters,echo=FALSE}
binningParameters(binalysis)
```


### Results   

The plots and tables below give an overview of the results of the spectral binning approach applied to this data set.



```{r rawFeaturesTable,echo=FALSE}
library(knitr)
rawFeat <- binalysis %>%
  binnedData() %>%
  map(~{
    d <- .
    d %>%
      rowid_to_column(var = 'Sample') %>%
      gather('Feature','Intensity',-Sample)
    }) %>%
  bind_rows() %>%
  mutate(Mode = str_sub(Feature,1,1)) %>%
  group_by(Mode) %>%
  summarise(`Number of bins` = n_distinct(Feature),`Missing Data (%)` = round(length(which(Intensity == 0))/length(Intensity) * 100,2))

rawFeat$Mode[rawFeat$Mode == 'n'] = 'Negative'
rawFeat$Mode[rawFeat$Mode == 'p'] = 'Positive'

kable(rawFeat,caption = 'Table overview of spectral bins returned for each acqusition mode')
```



```{r chromatograms,warning = FALSE,echo=FALSE}
binneR::plotChromatogram(binalysis)
```



```{r fingerprint,warning = FALSE,echo=FALSE}
plotFingerprint(binalysis)
```



```{r PurityCentrality,echo=FALSE}
plotPurity(binalysis) + plotCentrality(binalysis)
```



```{r TICplot,echo=FALSE}
binneR::plotTIC(binalysis, by = 'injOrder', colour = 'block')
```   


```{r RSDplot,echo=FALSE}
metaboMisc::plotRSD(binalysis) %>%
  walk(print)
```   

      
## Analysis
      
### Overview
      
**Technique:** `metabolyseR` analysis

```{r analysisOverview,echo = FALSE}
print(analysis)
```

### Parameters

```{r analysisParamters,echo=FALSE}
analysis@parameters
```
    
### Pre-treatment

```{r unsupervisedPlots,echo = F,fig.width = 10}
metabolyseR::plotPCA(analysis,cls = 'class',ellipses = TRUE,legend = TRUE) + plotUnsupervisedRF(analysis,cls = 'class',ellipses = TRUE,title = 'Multidimensional scaling (MDS) -
unsupervised Random Forest',legend = TRUE)
```
      
```{r LDAplot,echo = F,fig.width = 10}
plotLDA(analysis,cls = 'class',ellipses = TRUE,legend = TRUE)
```
    
```{r supervisedRFplots,echo = F,fig.width = 10}
  plotSupervisedRF(analysis,cls = 'class',label = NULL,ellipses = TRUE,title = 'Multidimensional scaling (MDS) -
supervised Random Forest',legend = TRUE,ROC = TRUE)
```
      
```{r ptPurityCentrality,echo=FALSE}
histBins <- 30

preTreatedFeatures <- analysis %>%
  preTreatedData() %>%
  colnames()

purCent <- binalysis %>%
  accurateData() %>%
  mutate(Feature = str_c(polarity,mz)) %>%
  dplyr::select(polarity,Feature,purity,centrality) %>%
  gather('Measure','Value',-polarity,-Feature) %>%
  mutate(polarity = as.character(polarity)) %>%
  na.omit() %>%
  filter(Feature %in% preTreatedFeatures)
  
    
purCent$polarity[purCent$polarity == 'n'] <- 'Negative'
purCent$polarity[purCent$polarity == 'p'] <- 'Positive'
    
purPlot <- purCent %>%
  filter(Measure == 'purity') %>%
  ggplot(aes(x = Value)) +
  geom_histogram(fill = ptol_pal()(5)[2],colour = 'black',bins = histBins) +
  theme_bw() +
  facet_wrap(~polarity) +
  ggtitle('Bin Purity Distribution') +
  theme(plot.title = element_text(face = 'bold'),
        axis.title = element_text(face = 'bold')) +
  xlab('Purity Measure') +
  ylab('Frequency')
    
centPlot <- purCent %>%
filter(Measure == 'centrality') %>%
ggplot(aes(x = Value)) +
  geom_histogram(fill = ptol_pal()(5)[2],colour = 'black',bins = histBins) +
  theme_bw() +
  facet_wrap(~polarity) +
  ggtitle('Bin Centrality Distribution') +
  theme(plot.title = element_text(face = 'bold'),
        axis.title = element_text(face = 'bold')) +
  xlab('Centrality Measure') +
  ylab('Frequency')
    
purPlot + centPlot + plot_layout(ncol = 1)
```


### Modelling

```{r modellingOverview, cache = FALSE,results = 'asis'}
modellingRes <- analysis %>%
  modellingResults()

if (names(modellingRes)[1] == 'randomForest') {
modellingRes %>%
  .[[1]] %>%
  map(measures) %>%
  bind_rows() %>%
  mutate_if(is.numeric,round,digits = 3) %>%
  datatable(rownames = F,filter = 'top',caption = 'Table of model measures')
}
```

```{r explanatoryFeatureOverview,cache = FALSE}
threshold <- 0.05

if (is.list(modellingRes %>% .[[1]])) {
  explanFeat <- modellingRes %>%
    .[[1]] %>%
    map(explanatoryFeatures,threshold = threshold) %>%
    bind_rows() 
} else {
  explanFeat <- modellingRes %>%
    .[[1]] %>%
    explanatoryFeatures(threshold = threshold)
}

explanFeat %>%
  group_by(Response,Comparison) %>%
  summarise(`# Explanatory` = n()) %>%
  datatable(rownames = F,filter = 'top',caption = str_c('Overview of numbers of explanatory features (p < ',threshold,')'))
```

```{r explanatoryFeatures,cache = FALSE}
ef <- explanFeat %>%
  tbl_df() %>%
  mutate_if(is.numeric,round,digits = 3)

names(ef)[names(ef) == 'adjusted.p.value'] <- 'adjustedPvalue' 

if ('adjustedPvalue' %in% names(ef)) {
 ef <- ef  %>%
    arrange(adjustedPvalue) 
} else {
  ef <- ef %>%
    arrange(Value)
}

if (nrow(ef) > 10000) {
  ef %>%
    .[1:10000,] %>%
    datatable(rownames = F,filter = 'top',caption = str_c('Table of top 10000 explanatory features (p < ',threshold,')'))
} else {
  ef %>%
    datatable(rownames = F,filter = 'top',caption = str_c('Table of explanatory features (p < ',threshold,')'))
}
```

```{r explanatoryHeatMap,fig.height = 10}
if (nrow(ef) > 2) {
  if (length(unique(ef$Feature)) > 150) {
    featNames <- F
  } else {
    featNames <- T
  }
  
  if (is.list(modellingRes[[1]])) {
    mod <- modellingRes[[1]][[1]]
  } else {
    mod <- modellingRes[[1]] 
  }
  
  plotExplanatoryHeatmap(mod,featureNames = featNames)
}
```

### Correlations

```{r correlationTable,echo = F,cache = F}
cors <- analysis %>%
  correlationResults()
if (nrow(cors) > 10000) {
  cors <- cors %>%
    arrange(desc(r)) %>%
    .[1:10000,]
  caption <- 'Table of top 10000 feature correlations'
} else {
  caption <- 'Table of feature correlations'
}
datatable(cors %>%
          mutate_if(is.numeric,round,digits = 3),
          rownames = F,filter = 'top',caption = caption)
```


## Annotation
      
### Overview
                      
**Technique:** `MFassign` assignment

```{r assignmentOverview,echo = FALSE}
print(assignment)
```

### Parameters

```{r assignmentParamters,echo=FALSE}
assignment@parameters
```

### Assignments  

```{r assignmentTable,echo = FALSE,cache = FALSE}
datatable(assignment %>%
  summariseAssignment(),
          rownames = F,filter = "top",caption = "Summary table of molecular formula assignments")
```


## Session Info

```{r sessionInfo}
sessionInfo()
```

    
-----------
Generated by [metaboReports](https://github.com/jasenfinch/metaboReports) v0.8.4